\LARGE \bf Multi-dimensional Optimization of Gasoline-fueled Variable Pitch Multirotor Aircraft

### 2016-04-13 paper.tex
final small revisions

Designing a gas powered multirotor aircraft requires several design decisions. When considering the total flight time of a the aircraft, foremost of these design decisions are the choices of rotor, engine and gas tank size.  These design decisions along with several other aircraft parameters all have trade-offs that makes it difficult for a designer to simultaneously consider which trade-offs will allow for the longest flight time and/or greatest payload capacity.

This paper presents analysis and optimization methods for designing a gas powered multirotor aircraft.  The analysis focuses on modeling the engine and aerodynamic components with a limited consideration of aircraft structure.  These analyses are then used within multi-dimensional optimization to produce a vehicle design capable of long duration flight.

@@

Multirotor aircraft are useful for a variety of applications.  Because multirotors can fly vertically, they do not require landing strips.  They can carry sensor equipment for scientific or surveillance purposes. Multirotor use in package delivery is being studied by several large companies such as Google, Amazon and DHL \cite{Amazon2014}.

Incorporating a gas powered engine into a multirotor requires more then just exchanging the electric motors and batteries for an engine and a gas tank, respectively. The control of such a vehicle would also fundamentally change.  Electric multirotors are controlled by varying propeller rotational speed to produce varying thrust and torque, which in turn controls the attitude of the aircraft. Gas engines lack the ability to vary speed quickly enough to control the attitude of a multirotor aircraft. This means a fixed speed and variable pitch propeller is better suited for a gas powered vehicle.

Variable pitch mulitrotor aircraft are not as common as variable speed multirotors but they have been explored to some degree. Control of variable pitch multirotor aircraft was studied by Cutler \cite{Cutler2012}. There is a small, electric variable pitch quadcopter commercially available \cite{stingray2016} for performing stunts and inverted maneuvers.  There are also several hobbyist attempts at gas powered variable pitch quad-rotor (quads) \cite{diy2016}, \cite{hackaday2016}. In our knowledge there has been no attempt to use a gas powered variable pitch quad to maximize the flight time for carrying a relatively large payload for commercial applications.

@@

FAA requirements \cite{faa2016} specify that model aircraft and UAVs of any sort must weigh less then 55 lbs (25 kg). UAVs or model aircraft weighing more than 55 lbs require special airworthiness certificates and are much more heavily regulated.

@@

@@

\begin{figurehere}
	\begin{center}
		\includegraphics[width=.5\textwidth]{current_capabilities.png}
		\caption{\textit{Graphic showing capabilities of current platforms and the goal of this project. The data points shown are specifications given by manufacturers of popular commercial and consumer UAVs and is collected from \cite{drones2016}.}} 
		\label{fig:current_cap}
	\end{center}
\end{figurehere}

The goal of this project is to expand the flight envelope of multirotor aircraft.  Figure~\ref{fig:current_cap} demonstrates the current capabilities of consumer and commercial multirotors and depicts the goal of this project.  We believe that a multirotor that could carry larger payloads for longer periods of time would greatly expand the type and number of applications for these vehicles. A multirotor with this capability would fundamentally change the way multirotors are used in commercial applications.

@@

Data was also used to produce a model of the power output of the engine as a function of size. Figure~\ref{fig:power} shows the collected data for engine size and power.  Again, we initially used a linear least squares fit to the data as well as compared our fit to data produced by Barnard Microsystems \cite{barnardmiro2016}. However, both models were inadequate for our analysis because they contain a non-zero y-intercept.  A non-zero y-intercept would allow an infinitely small engine to produce power and thus would allow the aircraft to fly infinitely. To correct for this issue, we applied a quadratic fit to the data and forced it to intercept through the origin.

@@

Within the optimization there are two sets of design variables.  First, the design variables that influence the size of the vehicle: the rotor radius $R_{rot}$, the engine size $S_{eng}$, and the fuel capacity $f_{capacity}$.  Then, there are the design variables that are used to determine feasibility of the design: fuel burn rate $f_{rate}$, rotor pitch angle $\theta_{pitch}$, and propeller head speed $\omega_{rot}$. Number of blades per rotor and number of rotors were also initially treated as design variables. However, with the current aerodynamic model, the optimizer will create as many rotors with as many blades as possible to minimize the rotational speed required of each rotor to reduce the overall drag. Since adding more rotors would have un-modeled adverse affects on cost, complexity, and weight these variables are fixed as parameters in the analysis. Moving forward we will assume four sets of two blade rotors because this is a common configuration that would be reasonably simple to build.

%The flight time of the vehicle is only directly proportional to two of the design variables: the fuel capacity $C_{fuel}$ and the fuel burn rate $r_{fuel}$.  Once the fuel is all burned then the vehicle must land.  Through the constraints imposed on the optimization the flight time becomes indirectly a function of the rest of the design variables.

@@

The optimization analysis makes several assumptions and approximations that would need to be improved in order to better trust the solution.  Furthermore, the optimization chose a design from continuous variables when only discrete off-the-shelf choices are available.

Future work would include using better aerodynamic modeling and validation with testing, adding structural analysis to find the weight of the frame, and optimizing under uncertainty due to inaccuracies in analysis.

>>>>

Designing a gas powered multirotor aircraft requires several design decisions. When considering the total flight time of a the aircraft, foremost of these design decisions are the choices of rotor, engine and gas tank size.  These design decisions along with several other aircraft parameters all have trade-offs that makes it difficult for a designer to simultaneously consider which design choice will allow for the longest flight time and/or greatest payload capacity.

This paper presents analysis and optimization methods for designing a gas powered multirotor aircraft.  The analysis focuses on modeling the engine and aerodynamic components with a limited consideration of aircraft structure.  These analyses are then used within multi-dimensional optimization to produce a vehicle design capable of both long duration flight and lifting heavy payloads.

@@

Multirotor aircraft are useful for a variety of applications. This is because multirotors can hover, fly vertically, and fly horizontally. Further, they do not require landing strips and can carry sensor equipment for scientific or surveillance purposes. Multirotor use in package delivery is being considered by several large companies such as Google, Amazon and DHL \cite{Amazon2014}.

Incorporating a gas powered engine into a multirotor requires more then just exchanging the electric motors and batteries for an engine and a gas tank. The control of such a vehicle would also fundamentally change.  Electric multirotors are controlled by varying propeller rotational speed to produce varying thrust and torque, which in turn controls the attitude of the aircraft. Gas engines lack the ability to vary speed quickly enough to control the attitude of a multirotor aircraft. This means a fixed speed and variable pitch propeller is better suited for a gas powered vehicle.

Variable pitch mulitrotor aircraft are not as common as variable speed multirotors but they have been explored to some degree. Control of variable pitch multirotor aircraft was studied by Cutler, et al. \cite{Cutler2012}\cite{Cutler2012a} and Segui-Gasco, et al. \cite{Segui-Gasco2014}. There is a small, electric variable pitch quadcopter commercially available \cite{stingray2016} for performing stunts and inverted maneuvers.  There are also several hobbyist attempts at gas powered variable pitch quad-rotors (quads) \cite{diy2016}, \cite{hackaday2016}. In our knowledge there has been no attempt to use a gas powered variable pitch multirotor to maximize the flight time for carrying a relatively large payload for commercial applications.

@@

Federal Aviation Administration (FAA) requirements \cite{faa2016} specify that model aircraft and UAVs of any sort must weigh less then 55 lbs (25 kg). UAVs or model aircraft weighing more than 55 lbs require special airworthiness certificates and are much more heavily regulated.

@@

\begin{figure}
	\begin{center}
		\includegraphics[width=.5\textwidth]{current_capabilities.png}
		\caption{\textit{Graphic showing capabilities of current platforms and the goal of this project. The data points shown are specifications given by manufacturers of popular commercial and consumer UAVs and is collected from \cite{drones2016}.}} 
		\label{fig:current_cap}
	\end{center}
\end{figure}

@@

The goal of this project is to expand the flight envelope of multirotor aircraft.  Figure~\ref{fig:current_cap} demonstrates the current capabilities of consumer and commercial multirotors and depicts the goal of this project.  We believe that a multirotor that could carry larger payloads for longer periods of time would greatly expand the type and number of applications for these vehicles. A multirotor with this capability would also fundamentally change the way multirotors are used in commercial applications.

@@

Data was also used to produce a model of the power output of the engine as a function of size. Figure~\ref{fig:power} shows the collected data for engine size and power.  Again, we initially used a linear least squares fit to the data as well as compared our fit to data produced by Barnard Microsystems \cite{barnardmiro2016}. However, both models were inadequate for our analysis because they contain a non-zero y-intercept.  A non-zero y-intercept would allow an infinitely small engine to produce power and thus would allow the aircraft to have infinite flight time. To correct for this issue, we applied a quadratic fit to the data and forced it to intercept through the origin.

@@

Within the optimization there are two sets of design variables.  First, the design variables that influence the size of the vehicle: the rotor radius $R_{rot}$, the engine size $S_{eng}$, and the fuel capacity $f_{capacity}$.  Then, there are the design variables that are used to determine feasibility of the design: fuel burn rate $f_{rate}$, rotor pitch angle $\theta_{pitch}$, and propeller head speed $\omega_{rot}$. Number of blades per rotor and number of rotors were also initially treated as design variables. However, with the current aerodynamic model, the optimizer will create as many rotors with as many blades as possible to minimize the rotational speed required of each rotor to reduce the overall drag. Since adding more rotors would have un-modeled adverse affects on cost, complexity, and weight these variables are fixed as parameters in the analysis. Moving forward we will assume four sets of two blade rotors because this is a common configuration that would be reasonably simple to build. Finally, although the selection of off-the-shelf components is a fundamentally discrete choice, for simplification within the optimization each design variable is treated as a continuous variable.

@@

The optimization analysis makes several assumptions and approximations that would need to be improved in order to better trust the solution.  Furthermore, the optimization chose a design from a continuous space when only discrete off-the-shelf choices are available. Further selection of the components would be needed to build a quad through a discrete optimization algorithm or analyzing all designs close to the continuous optimum. Future work would also include using better aerodynamic modeling and validation with testing, adding structural analysis to find the weight of the frame, and optimizing under uncertainty due to inaccuracies in analysis.
